---
title: "Inpatient Charge Data 2016"
author: "Antonio Avila"
date: "April 6, 2019"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse); library(modelr); library(maps);
```

Begin by loading in the data

```{r load data}
med_data = read_csv("medicare_data.csv", guess_max = 112000)
real_names = names(med_data)
names(med_data) <- c("DRG", "ID", "Provider", "Address", "City", "State", "Zip", "HRR", "Discharges", "AvgCharge", "AvgTotalPmts", "AvgMedPmts" )
```
There seems to be a problem parsing the data. The variable "Total Discharges" doesn't read in a few of the observations correctly because they're value is above 1,000. The commas seem to be affecting the parsing of those particular observations. In addition, The charges and payments variables are being parsed in as character types instead of numeric (or doubles) because of the dollar sign. 

```{r fix parsing}

parse2num <- med_data %>% 
    select("AvgCharge":"AvgMedPmts") %>% 
    purrr::map(parse_number) %>% 
    as_tibble()

med_data <- med_data %>% 
  select(-("AvgCharge":"AvgMedPmts")) %>% 
  bind_cols(parse2num)

```

Fixed the parsing issue for the Total Discharges column by extending the number of rows the read_csv() function reads in to determine the type of column it is to 120,000 since the first occurrence of a value over 1,000 occurred at about the 117,00th row, thus fixing the problem. Secondly,  converted the Average dollar payment columns into numeric columns, dropping the dollar symbol and ensuring the values are of the numeric type. 


Having fixed the parsing issues, I can  begin cleaning the data a little. I will begin by separating the code and descriptions from the DRG column to shorten it. The DRG Code are unique to their descriptions so I will seperate the two. The code will be used for general analysis since it is compact, making it easier to display on graphics while the description will be kept in case I want to group and subset of the data based on a more general type of procedure, i.e. heart procedures, respiratory, etc. This type of grouping can be easily be done by looking for key words in their descriptions, whereas the code provides no clue on how to do this, making it more difficult to automate.

```{r Seperate DRG Code}

med_data <- med_data %>% 
  separate(DRG, c("DRG_Code", "DRG_Descr"), sep = 3) 

med_data$DRG_Descr = str_sub(med_data$DRG_Descr, 4)
```

Even though procedure are already categorized into groups via the DRG classification system, it may be worth exploring whether certain groups of procedures are more expensive than others; for example, heart related procedures could be more or less expensive than other types of procedures even though not all heart procedures have the same level of severity. 

```{r Heart vs other}

med_data <- med_data %>% 
  mutate(Heart = str_detect(DRG_Descr, "HEART"))


med_data %>% 
  group_by(Heart) %>% 
  summarise(mean_charge = mean(AvgCharge)) %>% 
  ggplot(aes(Heart, mean_charge)) +
    geom_bar(stat = "identity", aes(fill = Heart), show.legend = FALSE)


```

In this case, it turns out that Heart related procedures as a whole are not more expensive when compared to all others, which is a little unexpected. I would expect heart related procedures to be more expensive in general because it is a vital organ and any type of major procedures is sure to be invasive, causing the need to consult a specialist. The average charge being lower may be because there are many more procedures provided that may not be very severe nor expensive. May be worth it to take a look and confirm if this is correct. 


```{r heart only}

heart_only <- med_data %>% 
  filter(Heart == TRUE)

heart_only %>% 
  count(DRG_Code)

# Visualizing the counts of each heart related DRG designated procedure. 
heart_only %>% 
  ggplot(aes(DRG_Code)) +
  geom_bar(aes(fill = DRG_Code), show.legend = FALSE)

heart_only %>% 
  group_by(DRG_Code) %>% 
  summarise(mean_charge = mean(AvgCharge))

# Visualizing the Mean Charge for a heart-related procedure by its DRG Code
heart_only %>% 
  group_by(DRG_Code) %>% 
  summarise(mean_charge = mean(AvgCharge)) %>% 
  ggplot(aes(DRG_Code, mean_charge)) +
    geom_bar(stat = "identity", aes(fill = DRG_Code), show.legend = FALSE) +
    geom_hline(yintercept = mean(heart_only$AvgCharge), color = "red") +
    labs(title = "Mean Charge per type of DRG Procedre",x = "DRG Code", y = "Mean Charge") +
    annotate("text", max(heart_only$DRG_Code), mean(heart_only$AvgCharge), vjust = -1, label = "Mean")



# Visualizing the proportion of payments. Skew towards the less expensive procedures.
total_charge = sum(heart_only$AvgCharge)
heart_only %>% 
  group_by(DRG_Code) %>% 
  summarize(group_charge = sum(AvgCharge), perc_charge = group_charge / total_charge) %>% 
   ggplot(aes(DRG_Code, perc_charge)) +
    geom_bar(stat = "identity", aes(fill = DRG_Code), show.legend = FALSE) +
    geom_ref_line(v = 4.5, size = 2)  +
    labs(title = "Percentage of Overall Mean per type of DRG Procedure", x = "DRG Code", y = "Percent of Total Charge") 


filter(heart_only, DRG_Code %in% c("001", "002", "215", "268")) %>% select(DRG_Descr) %>% unique()
filter(heart_only, !(DRG_Code %in% c("001", "002", "215", "268"))) %>% select(DRG_Descr) %>% unique()
```

Looking at only the heart related procedures confirm my hypothesis. There are significantly more heart related procedures whose expense is significantly less than there are expensive ones. As a results, they account for a larger proportion of the overall average charge, driving it down and explaining why it was unexpectedly low. The procedures desginated by DRG Codes 001, 002, 215, and 268 correspond to an invasive procedure, be it a heart transplant, heart assist implant, or aortic assist procedure. The less expensive and more common procedures correspond to some variation of heart failure, thus not requiring surgery at the time of the initial diagnosis; meaning it could be some kind of initial consultation resulting in heart failure diagnosis and possibly needing an invasive procedure in the future depending on the gravity of the situation. 
One thing to note from this for future reference is that procedures designated as having a major complication or comorbidity (MCC) tend to be more expensive than their non-MCC counterpart, which makes sense. May be worth investigating this comparison for all types of procedures in the future as well as to which places in the country have more MCC procedures. 



```{r}
heart_only %>% 
  filter(DRG_Code == "001") %>% 
  ggplot(aes(State, fill = State)) +
    geom_bar(show.legend = FALSE) +
    coord_flip()

heart_only %>% 
  filter(DRG_Code == "001") %>% 
  group_by(State) %>% 
  summarize(mean = mean(AvgCharge)) %>% 
  ggplot(aes(State, mean, fill = State)) +
    geom_bar(stat = "identity", show.legend = FALSE) +
    coord_flip()

```
Out of curiosity, I decided to look into where most of the heart transplant with MCCs tend to occur. To no suprise, they are most commonly performed in the most populous states: California, Florida, Texas, and New York. Out of these four states, California has a significantly higher cost for that type of procedure whereas the other three states have a similar cost to other states, on average. I would think it is because California has a high cost of living compared to most other staets, but New York has a similar cost of living. Suprisingly, Pennsylvania's average cost is comparable to California even though it is has about a quarter of the population and had about half of the procedure occur. The final thing to note, if you are in need in of a heart transplant or implant and have some sort of chronic disease that could complicate the surgery, go to Maryland. It still isn't cheap, but it is cheaper than mostly everywhere else. Arizona is comparable but it's too hot and dry out there and that is the last thing you need after a major operation. 



```{r  heat map of us}

# to create a map of the U.S, need the goegraphical coordinates. Loading in pre-built data 
#with coordinates but run into a problem.the data frame containing the coordinates has the 
#full state names whereas teh medicare data has abberviated state names so cant join directly

states <- map_data("state")

statenames.df <- bind_cols(tibble(state.abb), tibble(state.name))
statenames.df$state.name <- str_to_lower(statenames.df$state.name)

states <- statenames.df %>% left_join(states, by = c("state.name" = "region"))

charges <- med_data %>% group_by(State) %>% summarize(mean_charge = mean(AvgCharge))
map_df <- left_join(states, charges, by = c("state.abb" = "State"))

ggplot(map_df, aes(long, lat, group = group)) +
  geom_polygon(aes(fill = mean_charge)) +
  geom_path() +
  scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey90")+
  coord_map() +
  labs(x = "Longitude", y = "Lattitude", title = "Average Charge per Procedure per State", fill = "Average Charge")

```




```{r per state plots, fig.height = 6, include = FALSE}

med_data %>% 
  group_by(State) %>% 
  summarise(mean_charge = mean(AvgCharge)) %>%  
  ggplot(aes(mean_charge, reorder(State, mean_charge))) + 
    geom_point() +
    labs(x = "Mean Total Charge", y = "State")




med_data %>% 
  group_by(State) %>% 
  summarise(mean_disch = mean(Discharges)) %>%  
  ggplot(aes(mean_disch, reorder(State, mean_disch))) + 
    geom_point() +
    labs(x = "Mean Discharges", y = "State")

# med_data %>% 
#   ggplot(aes(Discharges, AvgCharge)) +
#     geom_point(aes(color = State))
# 

```

























































